---
title: "FECR - Examples"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Preparing the field
## Data import

SimuNet's networks are simulated relying mostly on two aspects: 

*  a weighted adjacency matrix `Adj`:  
_filled with integers representing the number of times two nodes have been observed associated_  
*  and a sampling effort `total_scan`:  
_representing the number of times the edges has been sampled to fill in `Adj`_  
\

A user willing to run simulations with SimuNet can import an `Adj` matrix and `total_scan` sampling effort by:  

*  randomly generate an `Adj` matrix and `total_scan`:  
```{r random init}
set.seed(42)
library(FECR)

# for a random directed matrix
n <- 5;nodes <- letters[1:n];
total_scan <- round(runif(1,0,100))
total_scan
Adj <- matrix(data = 0L,nrow = n,
              ncol = n,
              dimnames = list(nodes,nodes)
)
Adj

```
