---
title: "FECR - Use examples"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
In this page we show of to perform FECR tests - Fecal Egg Count Reduction tests - from individual
values of EPG: egg per gram of feces/fecal sediment.   

Implemented methods refer to Cabaret & Berrag (2004), as well as our method that improve their
proposed best approach.    

We demonstrate the use of these methods on simulated data, as well as the toy example presented in 
this reference article.

# Use example on simulated data
## EPG simulation
We simulate EPG values (one per individual) by drawing from poisson distributions as follow:

```{r random init}
library(FECR)

set.seed(42)

# Generate random EPG values, in the simple case of equal sample size
T1 <- rpois(50,100);T1.mean <- mean(T1)
T2 <- rpois(50,10);T2.mean <- mean(T2)
C1 <- rpois(50,100);C1.mean <- mean(C1)
C2 <- rpois(50,90);C2.mean <- mean(C2)
```
## FECR calculations
### When sample size is the same
The `FECR()` function of this `FECR` package takes vectors of EPG values as arguments `T1`, `T2`,
`C1`, and `C2`.

From these values, you can specify a method between `"Kochapakdee"`, `"Dash"`, `"Coles"`,
`"Cabaret1"`, `"Cabaret2"`.

Note that some of these methods do not require all arguments: for these methods, you will need to
specify what EPG values are inputted through named arguments (e.g. `T2 = your.vector`), otherwise an error
message will be displayed.

The logical argument `compute.CI` allows for the computation of a confidence interval (95% by
default) for individual-based methods (cf. Cabaret & Berrag, 2004)

```{r simple FECR}
# Calculation of FECR values according to different methods
FECR(T1.mean,T2.mean,method = "Kochapakdee")
FECR(T1,T2,C1,C2,method = "Dash")
FECR(T2 = T2.mean,C2 = C2.mean,method = "Coles")
FECR(T2 = T2,C2 = C2,method = "Coles")
FECR(T1,T2,method = "Cabaret1")
FECR(T1,T2,C1,C2,method = "Cabaret2")
FECR(T1,T2,C1,C2,method = "Cabaret2",compute.CI = TRUE)
FECR(T1,T2,C1,C2,method = "MacIntosh1",compute.CI = TRUE)
```
### When sample size is uneven
The last included method, `"MacIntosh1"`, is an improvement of the `"Cabaret2"` that:    

*  depends on the arbitrary pair-matching of individuals in the control (C) and test (T) groups
*  cannot be applied to unequal sample-sizes without arbitrarily subset the larger group
    
```{r unequal samp FECR, error=TRUE}
# Generate random EPG values, in the case of unequal sample size
T1 <- rpois(60,100);T1.mean<- mean(T1)
T2 <- rpois(60,10);T2.mean<- mean(T2)
C1 <- rpois(50,100);C1.mean<- mean(C1)
C2 <- rpois(50,90);C2.mean<- mean(C2)

# Calculation of FECR values according to our sample-size unsensitive method
FECR(T1,T2,C1,C2,method = "Cabaret2",compute.CI = TRUE)
FECR(T1,T2,C1,C2,method = "MacIntosh1")
FECR(T1,T2,C1,C2,method = "MacIntosh1",compute.CI = TRUE)

```

The different aspects of this improvement are discussed further with demonstrations and examples
[here](our_improvements.html).