---
title: "Why improve the method from Cabaret & Berrag (2004)?"
output: 
  html_document: 
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
---
The motivation that led us to propose an improvement over the individual method proposed by the
authors is that we identified the two pitfalls mentioned in the previous paragraph, i.e. the
arbitrary pair-matching and requirement of an even sample-size.

Hereafter, we demonstrate the issue from the paper's own toy example.

## EPG values import
```{r toy exp}
library(FECR)

set.seed(42)

# Generate random EPG values, in the simple case of equal sample size
T1 <- c(450,1250,1700,900,2550,500,750,850,500)
T1.mean <- mean(T1)
T2 <- c(50,0,0,0,50,450,400,0,0)
T2.mean <- mean(T2)
C1 <- c(750,600,500,700,1000,700,300,1250,1050)
C1.mean <- mean(C1)
C2 <- c(540,400,1000,1750,350,600,900,550,450)
C2.mean <- mean(C2)
```

## FECR calculations
As expected, we find FECR values similar to what the authors report:
```{r toy FECR}
# Generate random EPG values, in the simple case of equal sample size
FECR(T1.mean,T2.mean,method = "Kochapakdee")            # 90% in the original paper
FECR(T1,T2,C1,C2,method = "Dash")                       # 90% in the original paper
FECR(T2 = T2.mean,C2 = C2.mean,method = "Coles")        # 86% in the original paper
FECR(T1,T2,method = "Cabaret1",compute.CI = TRUE)       # 83%(58-99%) in the original paper
FECR(T1,T2,C1,C2,method = "Cabaret2",compute.CI = TRUE) # 84%(51-97%) in the original paper

FECR(T1,T2,C1,C2,method = "MacIntosh1",compute.CI = TRUE)
```
Our proposed method leads to a somewhat comparable value, albeit significantly lowered. We explain
more on this hereafter.


## Demonstration of the pit-falls
### The implicit arbitrary pair-matching


The toy-example's values only lead to the reported FECR when inputted in their specific order.
Shuffling this order is enough to change the FECR value:
```{r example pitfall 1}
set.seed(1000)
# shuffling T and C
new.T <- sample(1:length(T1))
new.C <- sample(1:length(C1))

T1.new <- T1[new.T]
T2.new <- T2[new.T]
C1.new <- C1[new.C]
C2.new <- C2[new.C]

# Recalculating FECRs
FECR(T1.new,T2.new,C1.new,C2.new,method = "Cabaret2",compute.CI = TRUE)
FECR(T1.new,T2.new,C1.new,C2.new,method = "MacIntosh1",compute.CI = TRUE)
```
Note that shuffling obviously has not change EPG distribution in T and C groups:
```{r suffling effects 1}
summary(T1)
summary(T1.new)
summary(C2)
summary(C2.new)
```

And neither have methods based on group-averaged EPG
```{r suffling effects 2}
FECR(T1.new,T2.new,C1.new,C2.new,method = "Dash")                       # 90% in the original paper
FECR(T2 = T2.new,C2 = C2.new,method = "Coles")        # 86% in the original paper
```
More precisely, here is the bootstrapped distribution of the FECR value itself (not its confidence
interval), across several
reshuffling:

```{r pitfall 1 details}
library(ggplot2)
library(magrittr)

set.seed(9000)

# replicating the shuffling of T and C
FECR.shuffled <- replicate(
  n = 1000,
  simplify = FALSE,
  expr = {
    # shuffle
    new.T <- sample(1:length(T1))
    new.C <- sample(1:length(C1))
    
    T1.new <- T1[new.T]
    T2.new <- T2[new.T]
    C1.new <- C1[new.C]
    C2.new <- C2[new.C]

    #FECR calculations
    data.frame(
      method = c("Cabaret2","MacIntosh1"),
      FECR = c(
        FECR(T1.new,T2.new,C1.new,C2.new,method = "Cabaret2",compute.CI = FALSE),
        FECR(T1.new,T2.new,C1.new,C2.new,method = "MacIntosh1",compute.CI = FALSE)
      )
    )
  }
) %>% 
  do.call(rbind,.)
FECR.shuffled %>% 
  ggplot(aes(x = FECR,fill = method))+
  facet_grid(method~.,scales = "free")+
  geom_histogram(binwidth = 2,alpha = 80)+
  guides(fill = "none")+
  scale_x_continuous(limits = c(-1,101))+
  theme_bw()
```

We can see that, depending on how individuals in T and C are paired, Cabaret & Berrag's method leads
to a FECR drawn from a bimodal distribution.

Our method consistently leads to a single FECR value, close to the mean of this distribution:
```{r pitfall 1 avg}
subset(FECR.shuffled,method == "MacIntosh1") %>% .$FECR %>%  unique()
subset(FECR.shuffled,method == "Cabaret2") %>% .$FECR %>%  mean()
```

### Uneven sample sizes
Let's simulate upon this toy example an uneven sample size between the T and C groups by randomly removing 2 individuals from one, e.g. T.

Since the `"Cabaret2"` method relies on a summation across n individuals in both the T and C groups, this requires subsetting the larger group too by randomly removing 2 individuals. This isn't a requirement for our method, which will keep using all the individuals from C despite T having 2 individuals less.
```{r example pitfall 2,error=TRUE}
set.seed(9999)
# removing 2 individuals from T and C
new.T <- sample(1:length(T1),length(T1) - 2)
new.C <- sample(1:length(C1),length(C1) - 2)

T1.new <- T1[new.T]
T2.new <- T2[new.T]
C1.new <- C1[new.C]
C2.new <- C2[new.C]

# Recalculating FECRs
FECR(T1.new,T2.new,C1.new,C2.new,method = "Cabaret2",compute.CI = TRUE)
FECR(T1.new,T2.new,C1,C2,method = "Cabaret2",compute.CI = TRUE)
FECR(T1.new,T2.new,C1,C2,method = "MacIntosh1",compute.CI = TRUE)
```

Again, if this is replicated many times, FECRs are impacted differently across the two methods:
```{r pitfall 2 details}
set.seed(777)

# replicating the shuffling of T and C
FECR.minus.2 <- replicate(
  n = 1000,
  simplify = FALSE,
  expr = {
    # shuffle
    new.T <- sample(1:length(T1),length(T1) - 2)
    new.C <- sample(1:length(C1),length(C1) - 2)
    
    T1.new <- T1[new.T]
    T2.new <- T2[new.T]
    C1.new <- C1[new.C]
    C2.new <- C2[new.C]
    
    #FECR calculations
    data.frame(
      method = c("Cabaret2","MacIntosh1"),
      FECR = c(
        FECR(T1.new,T2.new,C1.new,C2.new,method = "Cabaret2",compute.CI = FALSE),
        FECR(T1.new,T2.new,C1,C2,method = "MacIntosh1",compute.CI = FALSE)
      )
    )
  }
) %>% 
  do.call(rbind,.)
FECR.minus.2 %>% 
  ggplot(aes(x = FECR,fill = method))+
  facet_grid(method~.,scales = "free")+
  geom_histogram(binwidth = 2,alpha = 80)+
  guides(fill = "none")+
  scale_x_continuous(limits = c(-1,101))+
  theme_bw()
```

Therefore, on top of using all the data available, our method also leads to more stable FECR values.

### What about confidence intervals?

The same processes of reshuffling or removing datapoints can be replicated to calculate the
stability of confidence intervals' boundaries:
```{r pitfall CI details}
set.seed(1234)

# replicating the shuffling of T and C
CI.shuffled <- replicate(
  n = 100,
  simplify = FALSE,
  expr = {
    # shuffle
    new.T <- sample(1:length(T1))
    new.C <- sample(1:length(C1))
    
    T1.new <- T1[new.T]
    T2.new <- T2[new.T]
    C1.new <- C1[new.C]
    C2.new <- C2[new.C]
    Cab <- FECR(T1.new,T2.new,C1.new,C2.new,method = "Cabaret2",
                          compute.CI = TRUE,pb = FALSE)
    Mac <- FECR(T1.new,T2.new,C1.new,C2.new,method = "MacIntosh1",
                          compute.CI = TRUE,pb = FALSE)

    #FECR calculations
    data.frame(
      method = c("Cabaret2","MacIntosh1"),
      FECR = c(Cab,Mac),
      lower = c(
        attr(Cab,"CI")[1],
        attr(Mac,"CI")[1]
      ),
      upper = c(
        attr(Cab,"CI")[2],
        attr(Mac,"CI")[2]
      )
    )
  }
) %>% 
  do.call(rbind,.) %>% 
  cbind(rep = factor(rep(1:100,each = 2)),.)
CI.shuffled %>% 
  ggplot(aes(method,FECR,colour = method,group = rep))+
  geom_linerange(aes(ymin = lower,ymax = upper),position = position_dodge(.9))+
  geom_point(position = position_dodge(.9))+
  guides(colour = "none")+
  theme_bw()
```

Our proposed method is designed to calculate a FECR value (and its confidence interval) that is
unique for a given set of EPG values in T and C groups. The only slight variations observed
regarding the boundaries of the confidence interval is attributable to the bootstrap's own random
resampling.

## What about multiple values per individual and intra-individual variability?
None of the methods mentioned so far nor our fix of the Cabaret2 method handle multiple samples per
individual. In this FECR package, we propose a non-parametric approach to include individual
variability in the FECRs calculation.

### Simulating data
Here we simulate values for individuals of groups T and C, with different sample size per individual, group and period.
```{r data sim Mac2}
library(data.table,quietly = TRUE)
library(truncnorm)

set.seed(2001)
# custom simulating function:
sim_epg_data <- function(n,mean,runif.boundaries) {
  lapply(
    1:n,
    function(x) {
      # runif determine a random sample size per individual, and set the epg's sd to twice this
      # value
      runif(1,runif.boundaries[1],runif.boundaries[2]) %>% 
        {rtruncnorm(.,mean = mean,sd = . * 2,a = 0)} %>% 
        round %>% as.integer
    }
  )
}

# generating the data (generate list-column and unnest them for code brievity)
T1 <- data.table(id = letters[9:18],
                 epg = sim_epg_data(n = 10,mean = 65,c(3,10)))[,.(epg = epg[[1]]),by = id]
T2 <- data.table(id = letters[9:18],
                 epg = sim_epg_data(n = 10,mean = 20,c(6,8)))[,.(epg = epg[[1]]),by = id]
C1 <- data.table(id = letters[1:8],
                 epg = sim_epg_data(n = 8,mean = 70,c(6,12)))[,.(epg = epg[[1]]),by = id]
C2 <- data.table(id = letters[1:8],
                 epg = sim_epg_data(n = 8,mean = 65,c(1,12)))[,.(epg = epg[[1]]),by = id]

```
Note that ultimately, T1,T2,C1, and C2 now are data.frames (long format) containing column id (referring to the individual
identity) and epg (the value per sample). In the future, we will implement ways for the user to specify the column names.

```{r data vis}
rbind(cbind(group = "Treated",gp = "T1",T1),cbind(group = "Treated",gp = "T2",T2),
      cbind(group = "Control",gp = "C1",C1),cbind(group = "Control",gp = "C2",C2)) %>% 
  ggplot(aes(gp,epg,colour = id,fill = id))+
  facet_grid(.~group,scales = "free_x")+
  geom_boxplot(alpha = 0.3)+
  geom_jitter(alpha = .2)+
  guides(colour = "none",fill = "none")+
  xlab("Group per period (before = 1, after = 2)")+
  theme_minimal()

```

### FECR values 
With such a dataset, one needs first to average epg values per individual (and possibly by
group-period) before using any of the previously mentioned:
```{r data sim FECR,error=TRUE}
rbind(cbind(group = "Treated",gp = "T1",T1),cbind(group = "Treated",gp = "T2",T2),
      cbind(group = "Control",gp = "C1",C1),cbind(group = "Control",gp = "C2",C2)) %>% 
  .[,.(epg = mean(epg),l = quantile(epg,.025),u = quantile(epg,.975)),by = .(id,group,gp)] %>% 
  ggplot(aes(gp,epg,colour = id,fill = id))+
  facet_grid(.~group,scales = "free_x")+
  geom_line(aes(group = id),alpha = 0.2,colour = "grey50",position = position_dodge(0.2))+
  geom_linerange(aes(ymin = l,ymax = u),position = position_dodge(0.2))+
  geom_point(alpha = .5,position = position_dodge(0.2))+
  guides(colour = "none",fill = "none")+
  xlab("Group per period (before = 1, after = 2)")+
  theme_minimal()

T1.avg <- T1[,.(epg = mean(epg)),by = id]$epg;T2.avg <- T2[,.(epg = mean(epg)),by = id]$epg
C1.avg <- C1[,.(epg = mean(epg)),by = id]$epg;C2.avg <- C2[,.(epg = mean(epg)),by = id]$epg

# returns an error because of the uneven sample size
FECR(T1.avg,T2.avg,C1.avg,C2.avg,method = "Cabaret2",compute.CI = TRUE)

FECR(T1.avg,T2.avg,C1.avg,C2.avg,method = "MacIntosh1",compute.CI = TRUE)

# expected FECR given the simulated epg means in the treated group:
1 - (20 / 65) /  (65 / 70)
```
We can see that, in this simulation, the FECR value after averaging per individual is lower and less
variable than what could be expected by looking at the raw data and inputted mean epgs. Also, note
that the confidence interval being so narrow, it does not contain the value 66% obtained through the
inputted means.
  
In comparison, this is how our last method compares:
```{r data FECR Mac2}
FECR(T1,T2,C1,C2,boot = 500,method = "MacIntosh2")

# expected FECR given the simulated epg means in the treated group:
1 - (20 / 65) /  (65 / 70)
```
This last method, while returning a mean FECR lower than the one from the previous method, correctly
reflects how variable individual epg data are in the dataset, by reprensenting the ditribution of
FECR value one could have obtained given the individual variability. Note that this time, the "true"
value 66% calculated through the group mean epgs is within the 95% interval.
